A comparison of design-based and model-based approaches for finite population spatial data.

### 2021-12-22 inst/manuscript/manuscript.tex
minor updates to include estimated standard errors

In each of the 36 simulation scenarios, there were 2000 independent
simulation trials. In each trial, IRS and GRTS samples were selected and
then design-based and model-based analyses were used to estimate
(design-based) or predict (model-based) the mean and construct 95\%
confidence (design-based) or 95\% prediction (model-based) intervals.
Then we recorded the bias, squared error, and interval coverage for all
sampling-analysis combinations. After all 2000 trials, we summarized the
long-run performance of the combinations by calculating average bias,
rMS(P)E (root-mean-squared error for the design-based approaches and
root-mean-squared-prediction error for the model-based approaches), and
the proportion of times the true mean is contained in its 95\% interval.
The IRS algorithm, IRS variance estimator, GRTS algorithm, and local
neighborhood variance estimator are available in the \texttt{spsurvey}
\textbf{\textsf{R}} package (Dumelle et al., 2021). FPBK is available in
the \texttt{sptotal} \textbf{\textsf{R}} package (Higham et al., 2021).

@@

Fig. \ref{fig:figeff} shows the relative rMS(P)E of the four sampling
analysis combinations using the random location layout with
``IRS-Design'' as the baseline. The relative rMS(P)E is defined as
\begin{equation*}
\frac{\text{rMS(P)E of sampling-analysis combination}}{\text{rMS(P)E of IRS-Design}},
\end{equation*} When there is no spatial covariance (Fig.
\ref{fig:figeff}, ``Prop DE: 0'' row), the four sampling-analysis
combinations have approximately equal rMS(P)E. So using the GRTS
algorithm or a model-based analysis does not result in much, if any,
loss in efficiency compared to IRS-Design when there is no spatial
covariance. When there is spatial covariance (Fig. \ref{fig:figeff},
``Prop DE: 0.5'' and ``Prop DE: 0.9'' rows), GRTS-Model tends to perform
best, followed by GRTS-Design, IRS-Model, and finally IRS-Design, though
the difference in relative rMS(P)E among GRTS-Model, GRTS-Design, and
IRS-Model is relatively small. As the strength of spatial covariance
increases, the gap in rMS(P)E between IRS-Design and the other
sampling-analysis combinations widens. Finally we note that when there
is spatial covariance, IRS-Model outperforms IRS-Design by a large
margin, suggesting that the poor design properties of IRS are largely
mitigated by the model-based analysis. These conclusions are similar to
those observed in the grid location layout, so we omit a grid location
layout figure here. Tables for rMS(P)E in all 36 simulation scenarios
are provided in the supporting information.

\begin{figure}
  \centering
  \includegraphics[width = 1\linewidth]{figures/rmspe_eff.jpeg}
  \caption{Relative rMS(P)E in the simulation study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
  \label{fig:figeff}
\end{figure}

We also studied 95\% interval coverage among the sampling-analysis
combinations. The design-based confidence intervals and model-based
prediction intervals were constructed using the normal distribution.
Justification for this comes from the asymptotic normality of means via
the Central Limit Theorem. Fig. \ref{fig:figconf} shows the 95\%
interval coverage for each of the four sampling-analysis combinations in
the random location layout. Within each scenario, the sampling-analysis
combinations tend to have fairly similar interval coverage. Coverage in
the normal response scenarios was usually near 95\%, while coverage in
the lognormal response scenarios varied from from 90\% to 95\% but
increased with the sample size. At a sample size of 200, all four
sampling-analysis combinations had approximately 95\% interval coverage
in both response scenarios for all dependent error proportions. These
conclusions are similar to those observed in the grid location layout,
so we omit a grid location layout figure here. Tables for interval
coverage in all 36 simulation scenarios are provided in the supporting
information.

>>>>

In each of the 36 simulation scenarios, there were 2000 independent
simulation trials. In each trial, IRS and GRTS samples were selected and
then design-based and model-based analyses were used to estimate
(design-based) or predict (model-based) the mean and construct 95\%
confidence (design-based) or 95\% prediction (model-based) intervals.
Then we recorded the bias, squared error, and interval coverage for all
sampling-analysis combinations. After all 2000 trials, we summarized the
long-run performance of the combinations by calculating average bias,
rMS(P)E (root-mean-squared error for the design-based approaches and
root-mean-squared-prediction error for the model-based approaches), and
the proportion of times the true mean is contained in its 95\%
confidence (design-based) or 95\% prediction (model-based) interval. The
95\% confidence intervals (design-based) and 95\% prediction intervals
(model-based) were constructed using the normal distribution.
Justification for this comes from the asymptotic normality of means via
the Central Limit Theorem (under some assumptions). The IRS algorithm,
IRS variance estimator, GRTS algorithm, and local neighborhood variance
estimator are available in the \texttt{spsurvey} \textbf{\textsf{R}}
package (Dumelle et al., 2021). FPBK is available in the
\texttt{sptotal} \textbf{\textsf{R}} package (Higham et al., 2021).

@@

Fig. \ref{fig:rmspe_eff} shows the relative rMS(P)E of the four sampling
analysis combinations using the random location layout with
``IRS-Design'' as the baseline. The relative rMS(P)E is defined as
\begin{equation*}
\frac{\text{rMS(P)E of sampling-analysis combination}}{\text{rMS(P)E of IRS-Design}},
\end{equation*} When there is no spatial covariance (Fig.
\ref{fig:rmspe_eff}, ``Prop DE: 0'' row), the four sampling-analysis
combinations have approximately equal rMS(P)E. So using the GRTS
algorithm or a model-based analysis does not result in much, if any,
loss in efficiency compared to IRS-Design when there is no spatial
covariance. When there is spatial covariance (Fig. \ref{fig:rmspe_eff},
``Prop DE: 0.5'' and ``Prop DE: 0.9'' rows), GRTS-Model tends to perform
best, followed by GRTS-Design, IRS-Model, and finally IRS-Design, though
the difference in relative rMS(P)E among GRTS-Model, GRTS-Design, and
IRS-Model is relatively small. As the strength of spatial covariance
increases, the gap in rMS(P)E between IRS-Design and the other
sampling-analysis combinations widens. Finally we note that when there
is spatial covariance, IRS-Model outperforms IRS-Design by a large
margin, suggesting that the poor design properties of IRS are largely
mitigated by the model-based analysis. These conclusions are similar to
those observed in the grid location layout, so we omit a grid location
layout figure here. Tables for rMS(P)E in all 36 simulation scenarios
are provided in the supporting information.

\begin{figure}
  \centering
  \includegraphics[width = 1\linewidth]{figures/rmspe_eff.jpeg}
  \caption{Relative rMS(P)E in the simulation study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
  \label{fig:rmspe_eff}
\end{figure}

Fig. \ref{fig:mse_eff} shows the relative mean standard errors (MStdE)
of the four sampling-analysis combinations using the random location
layout with ``IRS-Design'' as the baseline. The MStdE is defined as
\begin{equation*}
\frac{\text{MStdE of sampling-analysis combination}}{\text{MStdE of IRS-Design}},
\end{equation*} Many general takeaways regarding MStdE are similar to
general takeaways regarding rMS(P)E: there seems to be no benefit to
using IRS, even when there is no spatial covariance; as the strength of
spatial covariance increases, the gap in MStdE between IRS-Design and
the other sampling-analysis combinations widens; and IRS-Model
outperforms IRS-Design by a large margin. This is not surprising because
all sampling-analysis combinations had nearly zero average bias, thus
rMS(P)E is driven by the variance of the estimators (design-based) or
predictors (model-based). We do note that between GRTS-Design and
GRTS-Model, GRTS-Design had lower MStdE when there was no spatial
covariance or a medium amount of spatial covariance (Fig.
\ref{fig:mse_eff}, ``Prop DE: 0'' and ``Prop DE: 0.5'' rows) and
GRTS-Model had lower MStdE when there was a high amount of spatial
covariance Fig. \ref{fig:mse_eff}, ``Prop DE: 0.9'' row). These
conclusions are similar to those observed in the grid location layout,
so we omit a grid location layout figure here. Tables for MStdE in all
36 simulation scenarios are provided in the supporting information.

\begin{figure}
  \centering
  \includegraphics[width = 1\linewidth]{figures/mse_eff.jpeg}
  \caption{Relative standard errors in the simulation study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
  \label{fig:mse_eff}
\end{figure}

Fig. \ref{fig:figconf} shows the 95\% interval coverage for each of the
four sampling-analysis combinations in the random location layout.
Within each scenario, the sampling-analysis combinations tend to have
fairly similar interval coverage. Coverage in the normal response
scenarios was usually near 95\%, while coverage in the lognormal
response scenarios varied from from 90\% to 95\% but increased with the
sample size. At a sample size of 200, all four sampling-analysis
combinations had approximately 95\% interval coverage in both response
scenarios for all dependent error proportions. These conclusions are
similar to those observed in the grid location layout, so we omit a grid
location layout figure here. Tables for interval coverage in all 36
simulation scenarios are provided in the supporting information.
